---
title: "Module 6"
author: "Emily Yang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

## Challenge 1
```{r Challenge 1}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/Country-Data-2016.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d) # First 6 rows
summary(d) # summary of each of the columns
names(d) # column names
# median area. Area 3rd column 
median(d[,c(3)], na.rm=TRUE)
# median population size. Pop size 2nd column 
median(d[,c(2)], na.rm=TRUE)

d$density <- d$population/d$area #use $ or it won't be a new column 
d <- d[order(-d$density), ] #order highest to lowest
d[1:10, ] # first 10 of d

d <- d[order(d$density), ] #order lowest to highest
d[1:10, ]

#Separate out the largest 20 countries
d <- d[order(-d$area), ] # Order by area from largest to smallest
size <- d[1:20, ] #New object with only the first 20 
summary(size)
# median area
median(size$area, na.rm=TRUE)
# median population size
median(size$population, na.rm=TRUE)

#Separate out the countries starting A-F
new <- d[grep("^[A-F]", d$country), ] # grep() search for matches of a pattern. ^ = first word. A-F = capital A-F  
summary(new)
median(new$area, na.rm=TRUE)
# median population size
median(new$population, na.rm=TRUE)
```

## Challenge 2
```{r Challenge 2}
par(mfrow = c(2, 3)) # tells R that all the graphs go in one image and how many rows and columns of it. 
attach(d)
na.omit(d$population)
na.omit(d$area)
boxplot(d$population)
boxplot(log(d$population))
boxplot(d$area)
boxplot(log(d$area))
barplot(d$population)
barplot(d$area)
detach(d)
```

## Challenge 3

```{r Challenge 3}
par(mfrow = c(1, 2)) 
attach(d)
hist(log(population), freq = FALSE, col = "red", main = "Plot 1", xlab = "log(population size)",
    ylab = "density", ylim = c(0, 0.2)) 
    abline(v = mean(log(population), na.rm = TRUE), col = "blue")
    
hist(log(area), freq = FALSE, col = "blue", main = "Plot 2", xlab = "log(area)",
    ylab = "density", ylim = c(0, 0.2))
 abline(v = mean(log(area), na.rm = TRUE), col = "green")# Adding mean area line

 par(mfrow = c(1, 1))  # set up one panel and redraw the log(population) histogram
hist(log(population), freq = FALSE, col = "white", main = "My Plot with Mean and Density",
    xlab = "log(population size)", ylab = "density", ylim = c(0, 0.2))
abline(v = mean(log(population), na.rm = TRUE), col = "blue")
lines(density(log(population), na.rm = TRUE), col = "green")
detach(d)
```

## Challenge 4
```{r Challenge 4}
sort(table(d$govt_form), decreasing = TRUE)
```

## Challenge 5
```{r Challenge 5}
m <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
n <- read.csv(m, header = TRUE, stringsAsFactors = FALSE)
attach(n)
head(n)
summary(n)
boxplot(log(Body_mass_female_mean) ~ Family, n) # anything ~ anything tells which rows to compare


library(ggplot2)
p <- ggplot(data = n, aes(x = Family, y = log(Body_mass_female_mean)))  #define the variables. aes means aesthetics 
p <- p + geom_boxplot()  #graph them in a boxplot
p <- p + theme(axis.text.x = element_text(angle = 90))  #put x-axis names at 90deg
p <- p + ylab("log(Female Body Mass)")  #rename y-axis title
p  #show me the graph

detach(n)
```

## Challenge 6

Make a scatter plot and also learn to add labels when not in ggplot.<br> https://www.tutorialkart.com/r-tutorial/r-plot-x-y-labels/#gsc.tab=0
``` {r Challenge 6}
attach(n)
par(mfrow = c(1, 2))
plot(x = Body_mass_female_mean, y = Brain_Size_Female_Mean, xlab="Brain Size Female Mean", ylab="Body Mass Female Mean") 
plot(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean), xlab="Log Brain Size Female Mean", ylab="Log Body Mass Female Mean") 
detach(n)

p <- ggplot(data = n, aes(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean),
    color = factor(Family)))  # first, we build a plot object and color points by Family. Needs to be a factor if its a character. Factor group identical charaacters.
p <- p + xlab("log(Female Body Mass)") + ylab("log(Female Brain Size)")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())  # then we modify the legend
p  # and, finally, we plot the object
# look at each family to find intra family differences
p <- p + facet_wrap(~Family, ncol = 4) 
p <- p + theme(legend.position = "none")
p
p <- p + geom_smooth(method = "lm", fullrange = TRUE) # lm = linear model. full range=TRUE telling it to draw the line across the entire graph.
p
```

``` {r Challenge 7}

library(ggplot2)
attach(n)
p <- ggplot(data = n, aes(x = log(Body_mass_female_mean), y = log(MaxLongevity_m)))
p <- p + geom_point()
p <- p + geom_smooth(method = "lm")
p

aggregate(n$Body_mass_female_mean ~ n$Family, FUN = "mean", na.rm = TRUE)

library(dplyr)
s <- filter(n, Family == "Hominidae" & Mass_Dimorphism > 2)
head(s)
s <- arrange(n, Family, Genus, Body_mass_male_mean)  # rearranging a data frame...
head(s)
s <- select(n, Family, Genus, Body_mass_male_mean)  # selecting specific columns...
head(s)
s <- rename(n, Female_Mass = Body_mass_female_mean)
head(s$Female_Mass)  # renaming columns...
s <- mutate(n, Binomial = paste(Genus, Species, sep = " "))
head(s$Binomial)  # and adding new columns...

s <- summarise(d, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
    na.rm = TRUE))
s

s <-                                                             #to create dataframe "s"
  n %>%                                                          #take dataframe "n"
  group_by(Family) %>%                                           #Group it by Family
  summarise(avgF = mean(Body_mass_female_mean, na.rm=TRUE),      #And calculate mean male BM
            avgM = mean(Body_mass_male_mean, na.rm=TRUE))        #And mean female BM
s
detach(n)
```

## Challenge 8
Learn how to pipe<br>

Acccording to Kamilar & Cooper’s (2013) dataset, what is the average male and female size, and body mass dimorphism of my two main study species (vervet monkeys, Chlorocebus pygerythrus; and woolly monkeys, Lagothrix lagotricha)? Which has a larger average female body mass? Which is more sexually dimorphic?<br>
Compare the body size of my two main study taxa at the Family level (i.e., Cercopithecidae vs. Atelidae) by plotting (using {ggplot2}) the body mass of males and females and sexual dimorphism. If you can, make the Cercopithecid boxes green, and the Atelid boxes purple.
``` {r Challenge 8}
library(dplyr)
s <-                                                  #to create dataframe "s"
  n %>%                                                    #take dataframe "n"
group_by(Family) %>%                                        #Group it by Family
summarise(avgF = mean(Body_mass_female_mean, na.rm=TRUE),  #calculate mean male BM
avgM = mean(Body_mass_male_mean, na.rm=TRUE))        #And mean female BM
s

s <- n %>%
    mutate(Binomial = paste(Genus, Species, sep = " ")) %>%
    select(Binomial, Body_mass_female_mean, Body_mass_male_mean, Mass_Dimorphism) %>%
    group_by(Binomial) %>%
    summarise(avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
        na.rm = TRUE), avgBMD = mean(Mass_Dimorphism, na.rm = TRUE))
x <- factor(c("Family", "Binomial"))

```

## Challenge 8 Problems
Acccording to Kamilar & Cooper’s (2013) dataset, what is the average male and female size, and body mass dimorphism of my two main study species (vervet monkeys, Chlorocebus pygerythrus; and woolly monkeys, Lagothrix lagotricha)? Which has a larger average female body mass? Which is more sexually dimorphic?

``` {r Challenge 8 Problem 1}
#filter out the two binomials
s %>% filter(Binomial == "Chlorocebus pygerythrus"| Binomial == "Lagothrix lagotricha")
```
Woolly monkeys have larger average body size but vervet monkeys have more sexual dimorphism<br>
<br>
Compare the body size of my two main study taxa at the Family level (i.e., Cercopithecidae vs. Atelidae) by plotting (using {ggplot2}) the body mass of males and females and sexual dimorphism. If you can, make the Cercopithecid boxes green, and the Atelid boxes purple.
``` {r Challenge 8 Problem 2}
#filter out the two Families

```

